---
title: "Univariate"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library("gridExtra")
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Data Analysis
### Simple frequency and plots for categorical variables
```{r echo=FALSE}
mydata<-read.csv("./data/dataset_Facebook.csv")
summary(mydata)
mydata$Category<-as.factor(mydata$Category)
mydata$Type<-as.factor(mydata$Type)
mydata$Post.Month<-as.factor(mydata$Post.Month)
mydata$Post.Weekday<-as.factor(mydata$Post.Weekday)
mydata$Post.Hour<-as.factor(mydata$Post.Hour)
mydata$Paid<-as.factor(mydata$Paid)
mydata<-na.omit(mydata) 
```


```{r message=FALSE,warning=FALSE, results='asis'}
library(summarytools)
freq(mydata$Category, plain.ascii = FALSE, style = "rmarkdown")
```



```{r echo=TRUE}
library(ggplot2)

p1<-ggplot(mydata,mapping = aes(x = Category, fill = Category))+geom_bar()
p2<-ggplot(mydata,mapping = aes(x = Type, fill = Type))+geom_bar()


p3<-ggplot(mydata, aes(Post.Month))+
  geom_bar(fill = "blue") 
p4<-ggplot(mydata, aes(Post.Hour))+
  geom_bar(fill = "red") 

grid.arrange(p1, p2,p3,p4 ,nrow = 2,ncol=2)
```



## Some observations 

* Most of the posts are Category 1 () posts 

* Most posts are video posts followed by status posts. Not many video posts 

* October has the most posts

* Early morning (post 12 AM) and 10 AM to 2 PM seem to be the preferred post timings

## Questions arising from observing the graph above

1. Do the brand engagement related variables (share,like,comment) differ based on post category, type , month,time,day of the post ?

***


### Continuous variables

```{r results='asis'}

library(summarytools)

# descr is used to summarise the numerical variables in the data frame
descr(mydata,transpose = TRUE,stats = c("mean","max","sd","skewness","kurtosis"))

```



#### Histograms

```{r}

p1 <- ggplot(data = mydata, mapping = aes(x = comment))+ geom_histogram(binsize = 0.5)

p2 <- ggplot(data = mydata, mapping = aes(x = share))+ geom_histogram(bins = 10)

p3 <- ggplot(data = mydata, mapping = aes(x = like))+ geom_histogram(bins = 10)


p4 <- ggplot(data = mydata, mapping = aes(x = Lifetime.Post.Consumers))+ geom_histogram(bins = 10)

grid.arrange(p1, p2,p3,p4 ,nrow = 2,ncol=2)


```
#### Box Plots

```{r}
b1<-ggplot(mydata, aes(x="share",y=share)) + 
  geom_boxplot()
b2<-ggplot(mydata, aes(x="like",y=like)) + 
  geom_boxplot()
b3<-ggplot(mydata, aes(x="comment",y=comment)) + 
  geom_boxplot()
b4<-ggplot(mydata, aes(x="Lifetime.Post.Consumers",y=Lifetime.Post.Consumers)) + 
 geom_boxplot()+coord_cartesian(ylim=c(0,12000))+scale_y_continuous(breaks=seq(0,12000,1000))
grid.arrange(b1, b2,b3,b4 ,nrow = 2,ncol=2)

```

##### Summary

The histograms show that all the continuous variables are right-skewed.Most of the values are less than the mean val.

1.Comments:Most of the posts have received close to 0 comments or have less than  hundred. There are a couple of posts which have comments more than 100.
2.All the other brand engagement metrics(share,like and Lifetime.Post.Consumers) show a similar pattern.A box plot would helps us identify the outliers.
Let's check this out next.

```{r}
a22 <- ggplot(mydata, aes(x = Category, y = comment)) +
  geom_boxplot()
a23 <- ggplot(mydata, aes(x = Type, y = comment)) +
  geom_boxplot()
a24 <- ggplot(mydata, aes(x = Paid, y = comment)) +
  geom_boxplot()
a25 <- ggplot(mydata, aes(x = Post.Hour, y = comment)) +
  geom_boxplot()
grid.arrange(a22, a23,a24,a25,nrow = 2,ncol=2)

```


##### Boxplots
Plot1:
Category 2(offer related posts) posts have generated the outlier.
Nevertheless, we can not that contest and discounts will generate more comments.
This should be verified by other relevant statistical tests.e.g. ANOVA

***

> Probability density Plots

```{r}
p1 <- ggplot(data = mydata, mapping = aes(x = comment))+ geom_density()


p2 <- ggplot(data = mydata, mapping = aes(x = share))+ geom_density()

p3 <- ggplot(data = mydata, mapping = aes(x = like))+ geom_density()


p4 <- ggplot(data = mydata, mapping = aes(x = Lifetime.Post.Consumers))+ geom_density()
p5 <- ggplot(data = mydata, mapping = aes(x = Lifetime.Post.Consumers))+ geom_density(alpha=0.3)

grid.arrange(p1, p2,p5,p4 ,nrow = 2,ncol=2)

```

