---
title: "Wine"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(png) 
```

```{r mystuff}
#'Reset environment before starting
set.seed(NULL)
rm(list = ls(all.names = TRUE))
#factoextra : to determine the optimal number clusters data visualization.
#fpc: cluster validation with silhouette width
library(factoextra)
library(fpc)
#'Load the data
# Read the contents of the csv file into a variable called wine_original
# "header=TRUE" implies that header is present in the data file

wine_original <- read.csv("./data/wines.csv", header = TRUE)
#'Examine the dataset

#str shows the structure of the object :variable type
str(wine_original)
#summary produces descriptive stats for the variables in the data
summary(wine_original)
#'Preprocessing
#We do not want to cluster region in the last  column
#Remove this and create a new dataframe to work with

wine.df<- wine_original[ , -c(14)]  # Drop region columns.
#Z-transform
wine.scaled.df<-as.data.frame(scale(wine.df))

#'Set the seed 
#Refer : http://rfunction.com/archives/62 K-means algorithm uses random number
#simulations to generate solutions.
# When we use any algorithm that uses random #number to simulate 
# we have to set the seed in r environment to get reproducible results
set.seed(96743)

#'Number of clusters
#To identify number of clusters apriori
# Use wss and silhouette metrics


#Plots Total within-clusters sum of squares vs. Number of clusters
fviz_nbclust(wine.scaled.df, kmeans, method = "wss")

#'The plot shows the elbow bend at **3 **
#Beyond this bend there is no value obtained in terms of wss
#Plots silhouette

fviz_nbclust(wine.scaled.df, kmeans, method = "silhouette")

#'The plot shows the peak at **3 ** again
#'Run k-means with 3 clusters
wine.seg <- kmeans(wine.scaled.df,centers = 3)
wine.seg

#'The results show the following
# 1. The number of clusters and the number of rows in each cluster
# 2. The means of the variables in each cluster
# 3. Clustering vector: A vector of integers that shows the cluster each observation belongs to

# Cluster size
wine.seg$size

# Cluster means
wine.seg$centers

#'Visualise the clusters

fviz_cluster(wine.seg, data = wine.scaled.df)

#'Tabulate the clusters and Profile
```{r, fig.width = 2}
# Small fig.width
#This is an excel screenshot to see the clusters clearly
img1_path <- "wine.png"
knitr::include_graphics(img1_path)

```

```{r mystuff2}


#Each row represents a cluster
#Get the max value loaded into each cluster
# Row 1 - Color intensity-Malic acid
# Row 2 - High in Proline/ Flavnoids-Low in ash akanity and non-flavnoid
# Row 3 - CDilute


mycluster_profile<-data.frame(
  cbind(wine_original,"cluster_number"=wine.seg$cluster))
# Add the names we gave for the clusters

mycluster_profile$cluster_name[mycluster_profile$cluster_number==1] <- "Color intensity-Malic acid"
mycluster_profile$cluster_name[mycluster_profile$cluster_number==2] <- "Proline"
mycluster_profile$cluster_name[mycluster_profile$cluster_number==3] <- "Dilute"

#Write the file to excel for further analysis
write.csv(mycluster_profile,file = "cluster_profiles.csv")



 
```

